Avoiding global deforestation by taxing land in agricultural production: The implications for global markets

### 2024-03-09 paper.Rmd
Update paper.Rmd

Both the GTAP and GTM models represent widely used models for policy analysis. In the case of GTAP, the model is used to describe the implications of various policy instruments for national and global markets, including prices, output, and trade. The GTM model is, on the other hand, capable to translating the impacts of population and global growth, and the changes in land returns, to changes in global forest stocks and calculate the impacts on total carbon contained in forests.

Even though the GTAP and GTM models capture important aspects of the global economy and the biophysical world, they do not model certain economic responses, and, instead, consider several important variables exogenous. In the case of GTAP, the model assumes that the amount of agricultural land is fixed, while the GTM considers the agricultural returns to land fixed. By combining the two models, we can fill each model's missing behavior for a more complete description of the global impacts.

The combination of the two models means that we use the GTAP model to create a link between the amount of available land and its returns for the benefit of the GTM model, which uses returns as an exogenous input, while at the same time we use the GTM model to create the response of  agricultural land quantity to returns in land returns for the benefit of the GTAP model, which accepts change in available agricultural land as an exogenous input. Specifically, we  solve both models concurrently for variable `qo` for land in the GTAP model and the parameter `RENTA` in the GTM model, which are internally consistent, i.e., the resulting rent change from agricultural land changes in the GTAP model are exactly the same as the changes in rents and the associated changes in agricultural land (calculated as the non-forested area from the total available land). Solving both models together in this way therefore allows us to expand their descriptive power.

We achieve a combined solution of GTAP and GTM by passing outputs back and forth between the models until a dual-state equilibrium is achieved. Establishing a connection between the GTAP model and the GTM model requires that we  model the return on land using the GTAP model and feed those results into the GTM model as exogenous inputs. The GTM model then endogenously determines the amount of land available for crop and livestock production, and we feed that back into the GTAP model as an exogenous input.

On the software level, we establish the connection between the models using an R script, which is capable of handling other software, including generating required inputs, executing it and reading in any outputs. This is required to operate the GTAP model, which is written in GEMPACK, and the GTM model, which is written in GAMS. As GEMPACK and GAMS were developed long before software integration became important or even possible, we use much more modern technologies made available through the R software [@R2021] to create the user functions that execute both models, collect their outputs, and automatically create any required input files.

>>>>

The GTAP and GTM models are both widely used for policy analysis. The GTAP model is frequently used to describe the implications of various policy instruments on national and global markets and their prices, output, and trade. The GTM model is often used to estimate the impacts of changes in population, GDP, and land rents on global forest stocks and carbon sequestration. While both models capture important aspects of the global economy and the biophysical world, they do not model certain economic responses, as both consider several important variables exogenous. By combining the two models, we expand each model's descriptive power and thus produce a better depiction of the global economy, which should allow for a more precise assessment of the impact of policy actions.

Specifically, we use the GTAP model to detail the impact on land rents based on the amount of available agricultural land. This benefits the GTM model, which uses returns as an exogenous input. We then use the GTM model to determine the change in agricultural land quantity based on the returns to land. This benefits the GTAP model, which takes the change in the amount of land available for crop and livestock production as an exogenous input. Using these links, we are able to solve both models concurrently, and outputs are passed back and forth between the models until we achieve a dual-state equilibrium. In the GTAP model, we solve for the variable `qo` for land, and in the GTM model, we solve for the parameter `RENTA`. These are internally consistent, i.e., the resulting rent change from agricultural land changes in the GTAP model mirrors the associated change in agricultural land (calculated as the non-forested area from the total available land) from the change in rents in GTM.

On the software level, we establish the connection between the models using R software [@R2021]. This is required as the two models were written using using different languages. Moreover, GTAP, which was written in GEMPACK, and GTM, which was written in GAMS, were developed long before software integration became important or even possible, and so we use an R script, which is capable of handling other software, including generating required inputs, executing them, and reading in any outputs, to create the user functions that execute both models, collect their outputs, and automatically create any required input files.
